Surrogate model of aligned-spin binary neutron star waveforms using Gaussian process regression

### 2018-01-16 paper/paper.tex
Add PhenomP ROQ paper reference

This work is a continuation of the surrogate developed in Ref.~\cite{LackeyBernuzziGalley2017} for a nonspinning BNS EOB waveform using the description for the point-particle and tidal terms by Bernuzzi {\it et al}.~\cite{Bernuzzi:2014owa}. Whereas in Ref.~\cite{LackeyBernuzziGalley2017} we constructed a time-domain surrogate with the goal of reproducing the original waveform model as accurately as possible, in this work, we construct a frequency-domain model which significantly speeds up parameter estimation by not requiring an online Fourier transform for each waveform evaluation. \MP{Is it really true that this is a significant speedup? FFT may still be cheaper than surrogate waveform generation.} It also enables additional techniques to accelerate likelihood evaluations such as reduced order quadrature~\cite{Antil2013, CanizaresFieldGair2013, CanizaresFieldGair2015} \MP{TODO: add citation to PhenomP ROQ paper here and later} and multi-banding~\cite{VinciguerraVeitchMandel2017}. The resulting surrogate is the first EOB waveform model that consistently incorporates spin and tides and is fast enough for production parameter estimation codes.

@@

The evaluation time has a flat cost of $\sim 0.02$~s to perform the GPR interpolation at each node. The rest of the time is spent resampling the waveform with spline interpolation. For a starting frequency of 10~Hz, this takes 0.3~s when the waveform is matched with data uniformly sampled at 4096Hz. For a starting frequency of 30~Hz, the total evaluation time is 0.04~s. For current parameter estimation codes, this is sufficient. However, one could further improve run times using reduced order quadrature~\cite{Antil2013, CanizaresFieldGair2013, CanizaresFieldGair2015} which requires frequency-domain waveforms, or multi-band waveform interpolation~\cite{VinciguerraVeitchMandel2017}. Finally, we note that the production MCMC sampler used for the GW170817 analysis has an autocorrelation length of $\mathcal{O}(10^4)$ for aligned-spin BNS models, so significant improvements to the parameter estimation runtime can be made through better samplers.

>>>>

This work is a continuation of the surrogate developed in Ref.~\cite{LackeyBernuzziGalley2017} for a nonspinning BNS EOB waveform using the description for the point-particle and tidal terms by Bernuzzi {\it et al}.~\cite{Bernuzzi:2014owa}. Whereas in Ref.~\cite{LackeyBernuzziGalley2017} we constructed a time-domain surrogate with the goal of reproducing the original waveform model as accurately as possible, in this work, we construct a frequency-domain model which significantly speeds up parameter estimation by not requiring an online Fourier transform for each waveform evaluation. \MP{Is it really true that this is a significant speedup? FFT may still be cheaper than surrogate waveform generation.} It also enables additional techniques to accelerate likelihood evaluations such as reduced order quadrature~\cite{Antil2013, CanizaresFieldGair2013, CanizaresFieldGair2015, Smith:2016qas} and multi-banding~\cite{VinciguerraVeitchMandel2017}. The resulting surrogate is the first EOB waveform model that consistently incorporates spin and tides and is fast enough for production parameter estimation codes.

@@

The evaluation time has a flat cost of $\sim 0.02$~s to perform the GPR interpolation at each node. The rest of the time is spent resampling the waveform with spline interpolation. For a starting frequency of 10~Hz, this takes 0.3~s when the waveform is matched with data uniformly sampled at 4096Hz. For a starting frequency of 30~Hz, the total evaluation time is 0.04~s. For current parameter estimation codes, this is sufficient. However, one could further improve run times using reduced order quadrature~\cite{Antil2013, CanizaresFieldGair2013, CanizaresFieldGair2015, Smith:2016qas} which requires frequency-domain waveforms, or multi-band waveform interpolation~\cite{VinciguerraVeitchMandel2017}. Finally, we note that the production MCMC sampler used for the GW170817 analysis has an autocorrelation length of $\mathcal{O}(10^4)$ for aligned-spin BNS models, so significant improvements to the parameter estimation runtime can be made through better samplers.
